---
title: "Assignment 6"
author: "Emil Jessen"
date: "19 okt 2020"
output:
  html_document: default
  pdf_document: default
---


```{r}
setwd("C:/Users/Lenovo/Desktop/Diverse dokumenter/Cognitive Science/cultural_data_science/assignment_6")
library("pacman")
p_load(gapminder, tidyverse)
```

Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}
df <- gapminder
df <- data.frame(lapply(df, as.character), stringsAsFactors=FALSE)


small_life_exp <- c()
medium_life_exp <- c()
large_life_exp <- c()

for (i in 1:length(df$country)){
  country <- df[i,"country"]
  if (substring(country,1,1) == "B"){
    if (df[i,"lifeExp"] <= 49){
      small <- df[i,"country"]
      small_life_exp <- c(small_life_exp,small)
    }
    else if (df[i,"lifeExp"] >= 49 & df[i,"lifeExp"] <= 70){
      medium <- df[i,"country"]
      medium_life_exp <- c(medium_life_exp,medium)
    }
    else {
      large <- df[i,"country"]
      large_life_exp <- c(large_life_exp,large)
    }
  }
}

# Printing each of the 3 groups
print(small_life_exp)
print(medium_life_exp)
print(large_life_exp)
```

Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.
```{r}
#Function
calc_GDP <- function(gdp_per_cap, pop){
  gdp <- (pop*gdp_per_cap)
  return(gdp)
}

#Making the data ready for the function
df$gdpPercap <- as.numeric(df$gdpPercap)
df$pop <- as.numeric(df$pop)
df <- as.data.frame(df)

#trying out the function, and here ascribing the values for a new column
df$gdp_country <- df %>% do(data.frame(df=calc_GDP(gdp_per_cap = df$gdpPercap, pop = df$pop)))

# Printing the gdp's for the years asked
print(df[df$country=="Denmark" & df$year == "1967", "gdp_country"])
print(df[df$country=="Denmark" & df$year == "1977", "gdp_country"])
print(df[df$country=="Denmark" & df$year == "1987", "gdp_country"])
print(df[df$country=="Denmark" & df$year == "1997", "gdp_country"])
print(df[df$country=="Denmark" & df$year == "2007", "gdp_country"])
```
