---
title: "SAFI_openrefine"
author: "Emil Jessen"
date: "17 sep 2020"
output:
  html_document: default
  pdf_document: default
---
1.
a) Good column names
b) No colour coding
c) Readme files
d) No comments, other than in a comment column
e) Missing data, use NA's, not zeros
f) One data frame per. document/sheet
g) Only 1-piece of information per block
h) Be considerate of formatting of dates
i) Consistency in variables

2.
No


3.
I did it in R, it is at the bottom, under "Which 2 month(s) are the most water-deprived/driest for the interviewed farmer households?"

4.
Described same place as 3.










#Answer to question 3, homework is at the bottom

```{r}
library(pacman)
setwd("C:/Users/Lenovo/Desktop/wd/cultural_data_science")
p_load(tidyverse, stringr)
safi <- read.csv("SAFI_openrefine.csv", header=T)
```

Cleaning up safi document
```{r}
unique(safi$village)

length(safi$village[safi$village == "Ruaca"])

safi$village[safi$village == "49"] <- NA
safi$village[safi$village == "Chirodzo"] <- "Chirdozo" 
safi$village[safi$village == "Ruca"] <- "Ruaca"
safi$village[safi$village == "Ruca-Nhamuenda"] <- "Ruaca - Nhamuenda"

```

How many unique days are spent in the team
```{r}
print(length(unique(safi$interview_date)))

# = 19
```

Which items_owned is the frequently recurring item?
```{r}
safi$items_owned_replaced <- gsub("\\[", "", safi$items_owned)
safi$items_owned_replaced <- gsub("\\]", "", safi$items_owned_replaced)
safi$items_owned_replaced <- gsub("\\;", "", safi$items_owned_replaced)
safi$items_owned_replaced <- gsub("\\'", "", safi$items_owned_replaced)
safi$items_owned_replaced <- gsub("   ", " ", safi$items_owned_replaced)
safi$items_owned_replaced <- strsplit(safi$items_owned_replaced, " ")
unique(unlist(safi$items_owned_replaced))
complete_list1 <- unlist(safi$items_owned_replaced)
complete_list1[complete_list1 == "" |complete_list1 == "NULL" ] <- NA
complete_list1 <- complete_list1[!is.na(complete_list1)]

print(table(complete_list1))


```

Which 2 month(s) are the most water-deprived/driest for the interviewed farmer households?
```{r}

# Here I take the entire dataframe, and remove the "[]".
View(safi$months_no_water)
safi$months_no_water_replaced <- gsub("\\[", "", safi$months_no_water)
safi$months_no_water_replaced <- gsub("\\]", "", safi$months_no_water_replaced)
safi$months_no_water_replaced <- gsub(" ; ", " ", safi$months_no_water_replaced)

# I then split the string into seperate elements (making each cell into a list)
safi$months_no_water_replaced <- strsplit(safi$months_no_water_replaced, " ")

# I then take each list for each cell and combine them into one long list
unique(unlist(safi$months_no_water_replaced))
complete_list <- unlist(safi$months_no_water_replaced)

# I remove NA's
complete_list[complete_list == "" |complete_list == "NULL" ] <- NA
complete_list <- complete_list[!is.na(complete_list)]

# Answer to question 3, homework:
# I then print the table, which has the count for each unique item in the list.
print(table(complete_list))

```



































